---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

```{r include = FALSE}
Sys.setlocale("LC_TIME", "C")
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r}
library(ggplot2);library(plyr);library(knitr)
unzip(zipfile = "activity.zip", exdir = ".", unzip = "internal")
df <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
df$date <- as.Date(df$date, "%Y-%m-%d")
str(df)
head(df, 3)
```

## What is mean total number of steps taken per day?

```{r}
df2 <- df[complete.cases(df),]
nr.steps <- aggregate(df2$steps, by=list(df2$date), sum)
ggplot(nr.steps, aes(x=nr.steps$x)) + geom_histogram(colour="black", fill="red",
binwidth = 5000) + 
ggtitle("Frequency of Number of Steps per Day") + 
labs(x = "Number of steps per day", y = "Frequency")
meansteps <- as.integer(mean(nr.steps$x))
mediansteps <- median(nr.steps$x)
```

The mean of total number of steps taken per day is **`r meansteps`** and the 
median is **`r mediansteps`**.

## What is the average daily activity pattern?

```{r}
activitypattern <- aggregate(df2$steps, by=list(df2$interval), mean)
ggplot(activitypattern, aes(x=activitypattern$Group.1)) + 
geom_line(aes(y=activitypattern$x)) + ggtitle("Daily Activity Pattern") + 
labs(x="5-minute interval", y="Number of steps per day")
max.nr.steps <- activitypattern[which.max(activitypattern$x),]$Group.1
```

The 5-minute interval containing the maximum number of 
steps is **`r max.nr.steps`**.

## Imputing missing values

```{r}
missing.values <- sum(rowSums(is.na(df)))
```

The total number of missing values is **`r missing.values`**.

Now, for filling in all of the missing values in the dataset, we will use 
the mean for the corresponding 5-minute interval:

```{r}
impute <- function(x, fun) {
  missing <- is.na(x)
  replace(x, missing, fun(x[!missing]))
}
df3 <- ddply(df, ~ interval, transform, steps = impute(steps, mean))
df3$date <- as.Date(df3$date, "%Y-%m-%d")
sum(is.na(df3))
```

We now have a new dataset df3 equal to the original dataset but with 
the missing data filled in.

And now for the histogram, mean and median:

```{r}
nr.steps2 <- aggregate(df3$steps, by=list(df3$date), sum)
ggplot(nr.steps2, aes(x=nr.steps2$x)) + geom_histogram(colour="black", 
fill="red", binwidth = 5000) + ggtitle("Frequency of Number of Steps per Day") +
labs(x = "Number of steps per day", y = "Frequency")
meansteps2 <- as.integer(mean(nr.steps2$x))
mediansteps2 <- as.integer(median(nr.steps2$x))
```

The mean of total number of steps taken per day is **`r meansteps2`** and 
the median is **`r mediansteps2`**.

**Mean value stays the same but there is a slight difference in median value.**

## Are there differences in activity patterns between weekdays and weekends?

First, we need to add one colum indicating wether a given date is a weekday 
or weekend day:

```{r}
df3$weekdayType <- ifelse(weekdays(df3$date) %in% c("Saturday", "Sunday"), 
"weekend", "weekday")
df3[sample(nrow(df3), 3),]
```

Now, the panel plot:

```{r}
activitypattern2 <- aggregate(df3$steps, by=list(df3$interval, df3$weekdayType),
mean)
ggplot(activitypattern2, aes(x=activitypattern2$Group.1)) + 
geom_line(aes(y=activitypattern2$x)) + facet_grid(Group.2~.) + 
ggtitle("Daily Activity Pattern") + labs(x="5-minute interval", 
y="Number of steps per day")
```